
Title: DSEA Data Visualization Workshop

Author: Kaijie(Lily)Fang; Yijin Du

Time: Oct.22 2021

##Introduction:ggplot2

Melt and Dcast 
```{r}
library(gapminder) 
library(reshape2)
library(tidyverse)

df <- select(gapminder, country, year, lifeExp)

#Long to wide format:
df_wide <- df %>% dcast(country~year)
df_wide
#Wide to long format:
df_long <- melt(df_wide, variable.name="year",
    value.name="lifeExp" )
df_long

```

Spread and Gather
```{r}
library(gapminder) 
library(tidyverse)
df <- select(gapminder, country, year, lifeExp)

#Long to wide format:
df_wide_1 <- df %>% spread(year, lifeExp)
df_wide_1
#Wide to long format:
df_long_1 <-  gather(df_wide, key = "year", value ="lifeExp", 2:13)
df_long_1
```

Example 1 
```{r}

library(tidyverse)
library(gapminder)
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point() + geom_smooth() 

```



## How to visualize different types of data with ggplot2

```{r}
##Install and load the packages (I have installed all the packages needed so I add "#" to not run it again. If you haven't installed these packages, delete the "#" and run the installation codes.)

#install.packages("ggplot2")
#install.packages("GGally")
#install.packages("ggExtra")
#install.packages("ggalluvial")
#install.packages("plotly")

# Please install it but do not load it for now.
#install.packages("cowplot")

library("ggplot2")
library("GGally")
library("ggExtra")
library("ggalluvial")
library("plotly")

# Load already installed packages
library("data.table")

ggplot2::ggplot()

```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import dataset from six countries: 
"Mexico", "Uruguay", "Japan", "Germany", "New Zealand", "Lebanon"
About the dataset: we will use the 2015 administration of the OECD’s Programme for International Student Assessment (PISA). PISA is a large-scale, international assessment that involves students, parents, teachers, and school principals from all over the world as participants. Every three years, PISA tests 15-year-old students from all over the world in reading, mathematics and science. The tests are designed to learn how well the students master key subjects in order to be prepared for real-life situations in the adult world. It also looks for the best practice.

```{r}
setwd("/Users/amanda/Desktop/TC/2021 fall/applied data/week3") #you need to change my file location to your file location where the dataset "pisa2015.csv" is located.
pisa <- fread("pisa2015.csv", na.strings = "")
```

Tidy
```{r}
#head(pisa)[1:10]
```

Transform
Convert a dichtomous item (yes/no) to numeric scoring
```{r}
bin.to.num <- function(x){
  if (is.na(x)) NA
  else if (x == "Yes") 1L
  else if (x == "No") 0L
}
```

create some variables
```{r}
pisa[, `:=` 
     (female = ifelse(ST004D01T == "Female", 1, 0),
       sex = ST004D01T,
       
       # At my house we have ...
       computer = sapply(ST011Q04TA, bin.to.num),
       software = sapply(ST011Q05TA, bin.to.num),
       internet = sapply(ST011Q06TA, bin.to.num))]

pisa[, `:=`
     (math = rowMeans(pisa[, c(paste0("PV", 1:10, "MATH"))], na.rm = TRUE),
       reading = rowMeans(pisa[, c(paste0("PV", 1:10, "READ"))], na.rm = TRUE),
       science = rowMeans(pisa[, c(paste0("PV", 1:10, "SCIE"))], na.rm = TRUE))]
```

Select some variables that we are interetsted in 
```{r}
country <- c("Mexico", "Japan","Germany", "Uruguay", "New Zealand", "Lebanon")
dat <- pisa[CNT %in% country,
             .(CNT, OECD, CNTSTUID, W_FSTUWT, sex, female,
              ST001D01T, computer, software, internet,
              ST011Q05TA, ST071Q02NA, ST071Q01NA, ST123Q02NA,
              ST082Q01NA, ST119Q01NA, ST119Q05NA, ANXTEST,
              COOPERATE, BELONG,  EMOSUPS, HOMESCH, ENTUSE,
              ICTHOME, ICTSCH, WEALTH, PARED, TMINS, ESCS,
              TEACHSUP, TDTEACH, IBTEACH, SCIEEFF,
              math, reading, science)
] #Variables and description: Table 5.1
```

create additional variables by recoding some of the existing variables
```{r}
dat <- dat[, `:=` (
  # New grade variable
  grade = (as.numeric(sapply(ST001D01T, function(x) {
    if(x=="Grade 7") "7"
    else if (x=="Grade 8") "8"
    else if (x=="Grade 9") "9"
    else if (x=="Grade 10") "10"
    else if (x=="Grade 11") "11"
    else if (x=="Grade 12") "12"
    else if (x=="Grade 13") NA_character_
    else if (x=="Ungraded") NA_character_}))),
  # Total learning time as hours
  learning = round(TMINS/60, 0),
  # Regions for selected countries
  Region = (sapply(CNT, function(x) {
    if(x %in% c("Mexico")) "N. America"
    else if (x %in% c( "Uruguay")) "S. America"
    else if (x %in% c("Japan")) "Asia"
    else if (x %in% c("Germany")) "Europe"
    else if (x %in% c( "New Zealand")) "Australia"
    else if (x %in% c("Lebanon")) "Middle-East"
  }))
)]
```

N count for the final dataset
```{r}
dat[,.N] # 35847 rows
```

Let's preview the final data
```{r}
head(dat)
```

Research Question 1:
To what extent was the six countries’ performance good or bad in PISA 2015?

Visualise
```{r}
#create a simple bar graph 
science_summary <- dat[, 
                       .(Science = mean(science, na.rm = TRUE), #average science scores by country
                         Freq = .N),
                       by = c("CNT")]
head(science_summary)
ggplot(data = science_summary,
       mapping = aes(x = CNT, y = Science)) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +
  labs(x = "country", y = "Science Scores") +
  theme_bw()
```


```{r}
#create a boxplot
ggplot(data = dat, mapping = aes(x = CNT, y = science)) +
  geom_boxplot() +
  labs(x=NULL, y="Science Scores") +
  theme_bw()

#switch x-axis and y-axis by using coord_flip() function
ggplot(data = dat,
       mapping = aes(x = CNT, y = science)) +
  geom_boxplot() +
  labs(x="Country", y="Science Scores", title = "Six Countries' Science Scores in PISA 2015") +
  coord_flip() +
  theme_bw()
```

```{r}
#show the mean values in boxplot
means <- dat[,
             .(science = mean(science)),
             by = CNT]

ggplot(data = dat,
       mapping = aes(x = reorder(CNT, science), y = science)) +
  geom_boxplot() +
  stat_summary(fun.y = mean, colour = "blue", geom = "point", 
               shape = 18, size = 3) + #mean score by countries
  labs(x="Country", y="Science Scores", title = "Six Countries' Science Scores in PISA 2015",
       subtitle = "Mexico, Japan, Germany, Uruguay, New Zealand, Lebanon") +
  coord_flip() +
  geom_hline(yintercept = 493, linetype="dashed", color = "red", size = 1) + #refence line-the average score across 6 countries
  theme_bw()
```
 
```{r}
#add some colors to our figure based on the region
ggplot(data = dat,
       mapping = aes(x = reorder(CNT, science), y = science, fill = Region)) +
  geom_boxplot() +
  stat_summary(fun.y = mean, colour = "blue", geom = "point", 
               shape = 18, size = 3) + #mean score by countries
  labs(x="Country", y="Science Scores", title = "Six Countries' Science Scores in PISA 2015",
       subtitle = "Mexico, Japan, Germany, Uruguay, New Zealand, Lebanon") +
  coord_flip() +
  geom_hline(yintercept = 493, linetype="dashed", color = "red", size = 1) +
  theme_bw()
```

```{r}
#science scores by gender and region
ggplot(data = dat,
       mapping = aes(x = science, fill = Region)) +
  geom_histogram(alpha = 0.5, bins = 50) +
  labs(x = "Science Scores", y = "Count",
       title = "Science Scores by Gender and Region") +
  facet_grid(. ~ sex) +
  theme_bw()

```

```{r}
#Create a correlation matrix plot
ggcorr(data = dat[,.(science, math, reading, ICTHOME, ESCS,
                     WEALTH, PARED, SCIEEFF, BELONG, grade)],
       method = c("pairwise.complete.obs", "pearson"),
       label = TRUE, label_size = 4)

```

Learning Time and Science Scores
```{r}
#Plotting the conditional (x-y) relationship of continuous variables
dat_small <- dat[,.SD[sample(.N, min(500,.N))], by = Region]

p1 <- ggplot(data = dat_small,
             mapping = aes(x = learning, y = science)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(x = "Weekly Learning Time", y = "Science Scores") +
  theme_bw()
p1
```

Learning Time and Science Scores
```{r}
#Plotting the conditional (x-y) relationship
p2 <- ggplot(data = dat_small,
             mapping = aes(x = learning, y = science,
                           colour = sex)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(x = "Weekly Learning Time", y = "Science Scores") +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.title = element_blank())
p2
```

Learning Time and Science Scores by Gender and Region
```{r}
#Plotting the conditional (x-y) relationship 
ggplot(data = dat_small,
       mapping = aes(x = learning, y = science, fill = Region)) +
  geom_point(aes(colour = sex)) +
  geom_smooth(method = "loess") +
  labs(x = "Weekly Learning Time", y = "Science Scores") +
  theme_bw()
```

Learning Time and Science Scores by Gender and Region
```{r}
ggplot(data = dat_small,
       mapping = aes(x = learning, y = science)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(x = "Weekly Learning Time", y = "Science Scores") +
  theme_bw() +
  theme(legend.title = element_blank()) +
  facet_grid(sex ~ Region)
```

Learning Time and Science Scores by Gender and Region
```{r}
#Plotting the conditional (x-y) relationship and marginal density of continuous variables
p2 <- ggplot(data = dat_small,
             mapping = aes(x = learning, y = science,
                           colour = sex)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(x = "Weekly Learning Time", y = "Science Scores") +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.title = element_blank())

ggMarginal(p2, type = "density", groupColour = TRUE, groupFill = TRUE)

```


Alluvial Plot
```{r}
dat_alluvial <- dat[, 
                    .(Freq = .N), 
                    by = c("Region", "sex", "ST123Q02NA")
                    ][,
                      ST123Q02NA := as.factor(ifelse(ST123Q02NA == "", "Missing", ST123Q02NA))
                      ]
levels(dat_alluvial$ST123Q02NA) <- c("Strongly disagree", "Disagree", "Agree",
                                     "Strongly agree", "Missing")

head(dat_alluvial)

ggplot(data = dat_alluvial,
       aes(axis1 = Region, axis2 = ST123Q02NA, y = Freq)) +
  scale_x_discrete(limits = c("Region", "Parents supporting\nachievement"),
                   expand = c(.1, .05)) +
  geom_alluvium(aes(fill = sex)) +
  geom_stratum() +
  geom_text(stat = "stratum", label.strata = TRUE) +
  labs(x = "Demographics", y = "Frequency", fill = "Gender") +
  theme_bw()

```


Create plot for color-blind people
```{r}
ggplot(data = dat_small,
       mapping = aes(x = learning, y = science, fill = Region)) +
  geom_point(aes(colour = sex)) +
  geom_smooth(method = "loess") +
  labs(x = "Weekly Learning Time", y = "Science Scores") +
  theme_bw()

cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

ggplot(data = dat_small,
       mapping = aes(x = learning, y = science, fill = Region)) +
  geom_point(aes(colour = sex) ) +
  geom_smooth(method = "loess") +
  labs(x = "Weekly Learning Time", y = "Science Scores") +
  theme_bw()+
  scale_fill_manual(values=cbPalette)+
  scale_colour_manual(values=cbPalette)

```

